---
title: "Analysis of environmental movements affectiveness"
subtitle: "INFO 526 - Summer 2024 - Final Project"
author: 
  - name: "Cole Johnson"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "This project will look into the effectiveness of the recent increase in environmental movements and how they have changed the pollutants being removed from rivers and streams"
format:
   html:
    code-tools: true
    code-overflow: wrap
    embed-resources: true
editor: visual
execute:
  warning: false
  echo: false
---

## Abstract

Add project abstract here.

## Introduction

Add project introduction here. Talk about the dataset

## Question 1: How are environmental movements affecting the Trash Wheels collections located that the end of rivers?

### Introduction
For this question I want to know the effects that environmental movements have on the pollutants that are being thrown into rivers where trash wheels are located. I would like to know if the amount of pollutants are decreasing and if the type of pollutants (recyclable versus non recyclable) are changing as a result of big environmental movements. I am interested in looking at this data because I have heard a ton of news over the last few years about environmental movements being more and more important and effective. I would like to know if all of these events actually change anything in peoples mind about pollutants or recyclables. 


### Approach

I first cleaned the dataset, making sure there were no blank data values that could change the analysis.

### Analysis

```{r setup}
if (!require("pacman")) 
  install.packages("pacman")

# install packages
pacman::p_load(tidyverse,
               glue,
               scales,
               ggthemes,
               viridis,
               openintro,
               patchwork,
               dsbox,
               ggridges,
               dplyr,
               fs,
               janitor,
               broom,
               colorspace,
               here,
               palmerpenguins,
               jpeg,
               ggpubr,
               tigris,
               sf,
               mapproj,
               geofacet,
               glue,
               ggrepel,
               ggspatial,
               rnaturalearth,
               rnaturalearthdata,
               readxl,
               patchwork,
               shiny,
               gghighlight
               ) 
#theme set
theme_set(theme_minimal(base_size = 14))
```

```{r}

trashwheel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-03-05/trashwheel.csv')

Movements <- tibble(date = c("2014-10-31", "2015-12-12", "2018-04-01", "2019-04-01","2021-01-01","2021-08-09"), description = c("IPCC Fifth Assessment Report", "Paris Agreement Signed", "Greta Thunberg starts school stike for climate", "First National CleanUp Day", "Biden rejoins Paris Accord", "IPCC Sixth Assessment Report"))

#fix dates

trashwheel <- trashwheel |>
  filter(!is.na(Name)) |>
  mutate(Date = paste0(str_split(Date, '\\/')[[1]][1],
                          '/',
                          str_split(Date, '\\/')[[1]][2]
                          )
         ) |>
  mutate(Date = paste(Date, Year, sep = '/')) |>
  mutate(Date = as.Date(Date, format = "%m/%d/%Y"))

Movements <- Movements |>
  mutate(date = as.Date(date)) |>
  mutate(datenew = date + 1)

#add recycle/nonrecycleqDate#add recycle/nonrecycleqDate#add recycle/nonrecyclequantity Cigarettebutts/100 because they are much smaller in size than everything else
trashwheel2 <- trashwheel |>
  mutate(Recyclable = PlasticBottles + GlassBottles + PlasticBags + Wrappers) |>
  mutate(NonRecyclable = Polystyrene + CigaretteButts/100 + SportsBalls) |>
  mutate(CigaretteButts = CigaretteButts/100)


#pivot_longer to make the data graphable
trashwheel3 <- trashwheel2 |>
  pivot_longer(
    cols = c("PlasticBottles", "Polystyrene", "CigaretteButts", "GlassBottles", "PlasticBags", "Wrappers", "SportsBalls", "Recyclable", "NonRecyclable"),
    names_to = "MaterialType",
    values_to = "Quantity"
  )


#Create Plot 1 - line graph with environmental movements annotated

p <- trashwheel3 |>
  ggplot() +
  geom_smooth(se = FALSE, aes(x = as.Date(Date), y = Quantity, color = MaterialType)) +
  geom_vline(data = Movements, 
             xintercept = as.Date(c("2014-10-31", "2015-12-12", "2018-04-01", "2019-04-01","2021-01-01","2021-08-09")),
             color = "black",
             size = 1
             ) +
  annotate("text", 
           x = as.Date(c("2014-7-31", "2015-9-12", "2018-1-01", "2019-1-01","2020-10-01","2021-05-09")),
           y = 4000,
           label = c("IPCC Fifth Assessment Report", "Paris Agreement Signed", "Greta Thunberg starts school stike for climate", "First National CleanUp Day", "Biden rejoins Paris Accord", "IPCC Sixth Assessment Report"),
           angle = 90,
           size = 3.5
           ) 

p1 <- p +
  labs(
    title = "Quantity of Pollutants Removed from all Trash Rivers",
    subtitle = "CigaretteButts quantity adjusted 100:1",
    caption = "Verticle lines represent major environmental movements
    Sources:tidyverse, various other sources",
    y = "Number of Pollutants",
    x = NULL
  ) +
  theme(
    plot.subtitle = element_text(size =  10)
  ) +
  guides(color = guide_legend(title = "Pollutant Type"))

  
p2 <- p +
  labs(
    title = "Quantity of Recyclable Pollutants Removed from all Trash Rivers",
    subtitle = "CigaretteButts quantity adjusted 100:1",
    caption = "Verticle lines represent major environmental movements
    Sources:tidyverse, various other sources",
    y = "Number of Pollutants",
    x = NULL
  ) +
  theme(
    plot.subtitle = element_text(size =  10)
  ) +   
  gghighlight(MaterialType == 'Recyclable') +
  guides(color = guide_legend(title = "Pollutant Type"))

p3 <- p +
  labs(
    title = "Quantity of Recyclable Pollutants Removed from all Trash Rivers",
    subtitle = "CigaretteButts quantity adjusted 100:1",
    caption = "Verticle lines represent major environmental movements
    Sources:tidyverse, various other sources",
    y = "Number of Pollutants",
    x = NULL
  ) +
  theme(
    plot.subtitle = element_text(size =  10)
  ) +   
  gghighlight(MaterialType == 'NonRecyclable') +
  guides(color = guide_legend(title = "Pollutant Type"))
 
p1
p2
p3
#Sources:
#Tidyverse
#https://www.ipcc.ch/site/assets/uploads/2018/04/ipcc_ar5_leaflet.pdf
#https://treaties.un.org/pages/ViewDetails.aspx?src=TREATY&mtdsg_no=XXVII-7-d&chapter=27&clang=_en
#https://www.euronews.com/green/2023/09/10/fridays-for-future-gretas-school-strikes-led-a-third-of-swiss-citizens-to-change-their-hab
#https://www.nationalcleanupday.org/earth-day
#https://apnews.com/article/joe-biden-executive-orders-b5b409da08e42414b9a12e2c67ee2df6
#https://www.ipcc.ch/assessment-report/ar6/
  
```
This visualization shows pollutant quantities being pulled from the end of rivers by Trash Wheels overtime. The visualization also shows the dates in which major environmental movements began or had a change. From the visualization we can see the non-recyclable quantity decrease fairly steadily over the entire period looked at. This visualization also tells us that certain items such as plastic bottles saw no change in quantities polluted into rivers over this period. Although this visualization shows the trend of quantities of pollutant types over time, I would like to see more in=depth of how each movement effected the quantities of recyclables and non-recyclables.

``` {r}
#manipulate data
trashwheel4 <- trashwheel3 |>
  mutate(
         FifthAssessmentStart = "2014-10-31", 
         FifthAssessmentEnd = "2015-12-12",
         ParisAgreementStart = "2015-12-12",
         ParisAgreementEnd = "2018-04-01",
         SchoolStrikeStart = "2018-04-01",
         SchoolStrikeEnd = "2019-04-01",
         FirstCleanUpDayStart = "2019-04-01",
         FirstCleanUpDayEnd = "2021-01-01",
         RejoinParisAccordStart = "2021-01-01",
         RejoinParisAccordEnd = "2021-08-09",
         SixthReportStart = "2021-08-09",
         SixthReportEnd = "2023-05-16"
         ) |>
  mutate(movement = case_when(
                              Date > FifthAssessmentStart & Date <= FifthAssessmentEnd ~ "IPCC Fifth Assessment Report",
                              Date > ParisAgreementStart & Date <= ParisAgreementEnd ~ "Paris Agreement Signed",
                              Date > SchoolStrikeStart & Date <= SchoolStrikeEnd ~ "Greta Thunberg starts school stike for climate",
                              Date > FirstCleanUpDayStart & Date <= FirstCleanUpDayEnd ~ "First National CleanUp Day",
                              Date > RejoinParisAccordStart & Date <= RejoinParisAccordEnd ~ "Biden Rejoins Paris Accord",
                              Date > SixthReportStart & Date <= SixthReportEnd ~ "IPCC Sixth Assessment Report",
                              TRUE ~ "FALSE"
                              
                              ),
         YearsEncompassed = case_when(
                              Date > FifthAssessmentStart & Date <= FifthAssessmentEnd ~ "1",
                              Date > ParisAgreementStart & Date <= ParisAgreementEnd ~ "2",
                              Date > SchoolStrikeStart & Date <= SchoolStrikeEnd ~ "1",
                              Date > FirstCleanUpDayStart & Date <= FirstCleanUpDayEnd ~ "2",
                              Date > RejoinParisAccordStart & Date <= RejoinParisAccordEnd ~ "1",
                              Date > SixthReportStart & Date <= SixthReportEnd ~ "2",
                              TRUE ~ "FALSE"
                              
                              )
         ) |>
  filter(movement != FALSE, MaterialType == c("NonRecyclable", "Recyclable"), !is.na(Quantity))

#Group by and create quantities

trashwheel5 <- trashwheel4 |>
  group_by(movement, MaterialType, YearsEncompassed) |>
  summarise(TotalQuantity = sum(Quantity)) |>
  ungroup() |>
  mutate(QuantityPerYear = TotalQuantity/as.numeric(YearsEncompassed))

trashwheel6 <- trashwheel5 |>
  group_by(movement) |>
  mutate(sum = sum(QuantityPerYear)) |>
  ungroup() |>
  mutate(Percentage = QuantityPerYear/sum) |>
  mutate(Percentage = round(Percentage, 2))

trashwheel6 |>
  ggplot(aes(x = movement, y = Percentage, fill = MaterialType)) +
  geom_bar(position = "fill", stat = "identity", color ='black', width = 0.9) +
  geom_text(aes(label = paste0(Percentage*100, "%")),
            position = position_stack(vjust = 0.5), size = 2) +
  scale_x_discrete(guide = guide_axis(n.dodge = 3),
                   limits = c("IPCC Fifth Assessment Report", "Paris Agreement Signed", "Greta Thunberg starts school stike for climate", "First National CleanUp Day", "Biden Rejoins Paris Accord", "IPCC Sixth Assessment Report")) +
  theme(axis.text.x = element_text(size = 10),
        plot.subtitle = element_text(size =  10),
        plot.title = element_text(size =  12)
        ) +
  labs(
    title = "Percentage of Recyclables Versus Nonrecyclables During Environmental Movements",
    subtitle = "Grouped by most recent environmental movement\nOrdered by Movement Date",
    caption = "Sources: Tidyverse, various other sources",
    y = NULL,
    x = "Environmenal Movements"
  ) +
  guides(fill = guide_legend(title = "Pollutant Type"))


#learned n.dodge from https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2

```
This visualization shows percentage of recyclable and non-recyclable pollutants in rivers following each environmental movement's occurrence. The visualization shows a direct comparison between recyclables and non-recyclable pollutants with percentage values of each which makes each environmental movements effect on types of pollutants very apparent. Although we see the same trend that was present in the prior visualization, the amount of non recyclables being greatly reduced over time, we also see something that wasn't very apparent in the last visualization. This visualization shows us that following the First National Cleanup Day, the percentage of non-recyclable pollutants actually increased in comparison to recyclable pollutants.


### Discussion

Talk about results here, what you gained from the graph

